Solution to ``Chapter 2: Basic tail and concentration bounds"

### 2020-08-13 skills/math/tail_bound/tailbound_solution.tex
tailbound

\begin{thm}[Chernoff bound]\label{thm:chernoff}
 Let $X$ be a random variable. Suppose that the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of 0; i.e., $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$. Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
	\begin{equation}\label{eq:chernoff1}
		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
	\end{equation}
	Optimizing the choice of $\lambda$ for the tightest bound, we obtain the Chernoff bound
	\begin{equation}\label{eq:chernoff2}
		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}.
	\end{equation}
\end{thm}

@@

	Plugging the inequality~\eqref{eq:e28tail} to $\bbE[Z] = \int_{0}^{+\infty} \bbP[Z \geq t] dt$, we have
	\begin{equation}
		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \colonequals I_1 + I_2.
	\end{equation}

	To evaluate $I_1$, we spilt the integral to avoid the minimization. Solving $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, the minimization term becomes
	\begin{align}
		\min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} = \begin{cases}
			\quad 1 & \text{when } t < 2\nu \sqrt{\log C},\\
			\quad C \exp \of{-\frac{t^2}{4 \nu^2}} & \text{when } t \geq 2\nu \sqrt{\log C}.
		\end{cases}
	\end{align}
	Therefore, we evaluate $I_1$ as follows,
	\begin{align}
		I_1 &= \int_{0}^{ 2\nu \sqrt{\log C}} 1 dt + \int_{2\nu \sqrt{\log C}}^{+\infty} C \exp \of{-\frac{t^2}{4 \nu^2}} dt\\
		(\text{let}\quad y = \frac{t}{2\nu} - \sqrt{\log C})\quad &=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
		&\leq  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2} dy.
	\end{align}
	By Gaussian integral $\int_{0}^{+\infty} e^{-x^2} dx = \frac{\sqrt{\pi}}{2}$, we obtain $I_1 \leq 2\nu (\sqrt{\pi} + \sqrt{\log C})$. Similarly, we evaluate $I_2$ as follows 
	\begin{align}
		I_2 &= \int_{0}^{4B \log C} 1 dt + \int_{4B \log C}^{+\infty} C \exp \of{-\frac{t}{4B}} dt\\
		 &= 4B (\log C + 1).
	\end{align}
	Hence, we obtain the expectation of $Z$,
	\begin{equation}
		\bbE[Z] = I_1 + I_2 \leq  2\nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (\log C + 1).
	\end{equation}

@@

\subsection{Exercise 2.9}\label{exe:29}
\begin{lem}[Sharp upper bounds on binomial tails]
	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$. Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$. The tail probability of $Z_n$ is upper bounded as
	\begin{equation}\label{eq:lembinupper}
		\bbP\off{Z_n \leq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha),
	\end{equation}
	where the quantity 
	\begin{equation}\label{eq:kldiv}
		D(\delta \Vert \alpha) \colonequals \delta \log \frac{\delta}{\alpha} + (1-\delta) \log \frac{1-\delta}{1-\alpha}
	\end{equation}
	is the KL divergence between the Bernoulli distributions with parameters $\delta$ and $\alpha$, respectively.

@@

\begin{proof}
    Applying the Chernoff inequality to the random variable $Z_n$, we have
    \begin{equation}\label{eq:e29upper}
    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda \in \bbR,
    \end{equation}
    where the second equality follows from the moment generating function of $Z_n$,
    \[\bbE\off{e^{\lambda Z_n}} = \of{1 - \alpha + \alpha e^{\lambda}}^n. \]
    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition. Plugging the minimizer into equation~\eqref{eq:e29upper}, we obtain the result
    \[ \bbP\off{Z_n \leq \delta n}  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha).  \]

@@

@@

	\vspace{0.2cm}
	Since $Z_n$ is a binomial variable with size $n$ and probability $\alpha$, we have
	\begin{align}
		\bbP[Z_n \leq n\delta ] &= \sum_{k = 1}^m \begin{pmatrix}
			n \\ k
		\end{pmatrix} \alpha^k (1-\alpha)^{n - k}\\
		& \geq \begin{pmatrix}
			n \\ m
		\end{pmatrix} \alpha^m (1-\alpha)^{n - m}. \label{eq:e210pro1}
	\end{align}
	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1} , we obtain
	\[ \frac{1}{n} \log \bbP[Z_n \leq n\delta ] \geq \frac{1}{n} \log \begin{pmatrix}
			n \\ m		\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha). \]

@@

	\vspace{0.2cm}
	Plugging equation~\eqref{eq:e2102} into equation~\eqref{eq:e2101}, we have
	\begin{equation}\label{eq:e210pro3}
		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \phi(\tilde \delta) + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha) - \frac{\log(n+1)}{n} \geq -D(\delta \Vert \alpha ) - \frac{\log(n+1)}{n}.
	\end{equation}
	Multiplying $n$ and exponentiating on the both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
	\begin{equation}
		\bbP \off{Z_n \leq n\delta} \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)}.
	\end{equation}
\end{proof}

@@

\subsection{Exercise 2.11}
\begin{lem}[Gaussian maxima]
Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d. normal random variables following $N(0,\sigma^2)$. Consider the random variable $Z_n \colonequals \max_{i\in[n]} |X_i|$. Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the expectation of $Z_n$ is upper bounded as follows
\begin{equation}\label{eq:e2111}
	\bbE[Z_n] \leq \sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}, \quad \text{for all } n\geq 2.
\end{equation}
\iffalse
The lower bound of $\bbE[Z_n] $ is
\begin{equation}\label{eq:e2112}
	\bbE[Z_n] \geq (1- 1/e) \sqrt{2\sigma^2 \log n} , \quad \text{for all }n \geq 5.
\end{equation}
\fi

@@

	\vspace{0.2cm}
	Consider the tail probability of $Z_n$. We have
	\begin{equation}
		\bbP[Z_n \geq t] = \bbP[\max_{i \in [n]} |X_i| \geq t] = 1 - \bbP[\max_{i \in [n]} |X_i| < t] = 1 - (1- \bbP[|X_1| \geq t])^n,
	\end{equation}
	where the last equality follows from the independence of $\{X_i\}_{i=1}^n$. By Bernoulli's inequality, we have
	\begin{equation}
		(1- \bbP[|X_1| \geq t])^n \geq 1 - n \bbP[|X_1| \geq t], \quad \text{for all } t >0.
	\end{equation} 
	Therefore, the expectation of $Z_n$ follows
	\begin{equation}\label{eq:e211pro1}
		\bbE[Z_n] = \int_{0}^{+\infty} \bbP[Z_n \geq t] dt \leq  c +  \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt, \quad \text{for all } c > 0.
	\end{equation}
	 Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the tail bound of $|X_1|$ satisfies
	 \begin{equation}\label{eq:e211proupper}
	 	\bbP[|X_1| \geq t] \leq 2 \sqrt{\frac{2}{\pi}} \frac{\sigma}{t} e^{-\frac{t^2}{2\sigma^2}}.
	 \end{equation}
	 Hence, the last integral in equation~\eqref{eq:e211pro1} follows
	 \begin{align}
	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\ (\text{let } u = \frac{ t}{\sigma}) \quad  
	 	 %& = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} e^{-\frac{u^2}{2}} dt \\
	 	 & \leq \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} u e^{-\frac{u^2}{2}} dt \\
	 	 & = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}  e^{-\frac{c^2}{2\sigma^2}}. \label{eq:e211pro2}
	 \end{align}
	 For all $n \geq 2$, let $c = \sqrt{2\sigma^2 \log n}$ and plug the $c$ into equations~\eqref{eq:e211pro1} and \eqref{eq:e211pro2}. Then, we obtain the result
	 \begin{equation}
	 	\bbE[Z_n] \leq \sqrt{2\sigma^2 \log n} + \frac{2\sigma}{\sqrt{2\log n}}   \sqrt{\frac{2}{\pi}}  \leq \sqrt{2\sigma^2 \log n} + \frac{4\sigma}{\sqrt{2\log n}}, \quad \text{for all }n \geq 2.
	 \end{equation}

	 \iffalse
	 \vspace{0.2cm}
	 Next, we prove the equation~\eqref{eq:e2112}.

@@

	 By Markov's inequality, the expectation of $Z_n$ satisfies
	 \begin{equation}
	 	\bbE[Z_n] \geq t \bbP[Z_n \geq t],\quad \text{for all }t \geq 0.
	 \end{equation}
	 By the independence of $\{X_i\}_{i=1}^n$, the tail probability of $Z_n$ is
	 \begin{equation}
	 	\bbP[Z_n \geq t] = 1 - (1 - \bbP(|X_1| \geq t))^n.
	 \end{equation}

@@

@@

@@

@@

@@

		\item[2.] Case 2: Suppose $t < 2\Delta$.
	\end{enumerate}
	\end{proof}

@@

    \subsection{Exercise 2.16}
    \begin{lem}
    	Let $\{X_i\}_{i=1}^n$ be a sequence of independent variables taking values in a Hilbert space $\mathbb{H}$. Suppose that $\onorm{X_i}_{\mathbb{H}} \leq b_i$ almost surely, for all $i \in [n]$. Consider the real valued random variable $S_n = \onorm{\sum_{i=1}^n X_i}_{\mathbb{H}}$. The concentration bound of $S_n$ is
    	\begin{equation}\label{eq:e216a}
    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all } \delta \geq 0,
    	\end{equation}
    	where $b^2 = \frac{1}{n} \sum_{i=1}^n b_i^2$. The upper tail probability bound of $S_n$ is
    	\begin{equation}\label{eq:e216b}
    		\bbP \off{\frac{S_n}{n} \geq a + \delta} \leq e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0,
    	\end{equation}
    	where $a = \sqrt{\frac{1}{n} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
    \end{lem}

@@

@@

    The proof for Hanson-Wright Inequality under the special case that $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.

@@

    	Consider $p_1$. For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and thus $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$. Given a constant $c$, $cX^2$ is still sub-exponential with parameters $(2 c, 4 c^2)$. Hence, for all $i \in [n]$, the variable $\mQ_{ii} X_i^2$ is sub-exponential with parameters $(2 \mQ_{ii}, 4 \mQ^2_{ii} )$. By the Bernstein-type inequalities(non-asymptotic 5.16), the probability $p_1$ is upper bounded as following,
    	\begin{equation}
    		p_1 = \bbP \off{\sum_{i=1}^n { \of{ \mQ_{ii}X^2_i - \mQ_{ii}\bbE[X^2_i]   } \geq \frac{t}{2} }} \leq \exp \of{ - \min \offf{  \frac{c t^2}{\sum_{i}^n \mQ^2_{ii} }, \frac{c't}{\max_{i \in [n]  }|\mQ_{ii}| } }},
    	\end{equation}
    	where $c$ and $c'$ are two constants independent with $n$ and $t$.

@@

@@

@@

@@

@@

 	 An alternative version of the proof when $\Delta_R > 0$ is below.

>>>>

\begin{thm}[Chernoff bound]\label{thm:chernoff}
 Let $X$ be a random variable. Suppose that the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of 0; i.e., $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$. Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
	\begin{equation}\label{eq:chernoff1}
		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (0,b).
	\end{equation}
	Optimizing the choice of $\lambda$ for the tightest bound, we obtain the Chernoff bound
	\begin{equation}\label{eq:chernoff2}
		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}.
	\end{equation}
\end{thm}

@@

	Plugging the inequality~\eqref{eq:e28tail} to $\bbE[Z] = \int_{0}^{+\infty} \bbP[Z \geq t] dt$, we have
	\begin{equation}
		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \equalscolon I_1 + I_2.
	\end{equation}

	To evaluate $I_1$, we solve $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, and the minimization term becomes
	\begin{align}
		\min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} = \begin{cases}
			\quad 1 & \text{when } t < 2\nu \sqrt{\log C},\\
			\quad C \exp \of{-\frac{t^2}{4 \nu^2}} & \text{when } t \geq 2\nu \sqrt{\log C}.
		\end{cases}
	\end{align}
	Therefore, let $ y = \frac{t}{2\nu} - \sqrt{\log C}$, we obtain $I_1$ ,
	\begin{align}
		I_1 &= \int_{0}^{ 2\nu \sqrt{\log C}} 1 dt + \int_{2\nu \sqrt{\log C}}^{+\infty} C \exp \of{-\frac{t^2}{4 \nu^2}} dt\\
		&=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
		&\leq  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2} dy.
	\end{align}
	By Gaussian integral $\int_{0}^{+\infty} e^{-x^2} dx = \frac{\sqrt{\pi}}{2}$, we obtain $I_1 \leq 2\nu (\sqrt{\pi} + \sqrt{\log C})$. Similarly, we evaluate $I_2$ as follows 
	\begin{align}
		I_2 &= \int_{0}^{4B \log C} 1 dt + \int_{4B \log C}^{+\infty} C \exp \of{-\frac{t}{4B}} dt\\
		 &= 4B (\log C + 1).
	\end{align}
	Hence, we obtain the expectation of $Z$,
	\begin{equation}
		\bbE[Z] = I_1 + I_2 \leq  2\nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (\log C + 1).
	\end{equation}

@@

\subsection{Exercise 2.9}\label{exe:29}
\begin{lem}[Sharp upper bounds on binomial tails]\label{lem:sharpbin}
	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$. Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$. The left tail probability of $Z_n$ is upper bounded as
	\begin{equation}\label{eq:lembinupper}
		\bbP\off{Z_n \leq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha),
	\end{equation}
	where the quantity 
	\begin{equation}\label{eq:kldiv}
		D(\delta \Vert \alpha) \colonequals \delta \log \frac{\delta}{\alpha} + (1-\delta) \log \frac{1-\delta}{1-\alpha}
	\end{equation}
	is the KL divergence between the Bernoulli distributions with parameters $\delta$ and $\alpha$, respectively.

@@

\begin{proof}
    Applying the Chernoff inequality to the random variable $-Z_n$, we have

    \begin{equation}
    	\bbP[-Z_n \geq -\delta n] =  \bbP[Z_n \leq \delta n]  \leq  \frac{\bbE\off{e^{-\lambda Z_n}}}{e^{-\lambda \delta n}}, \quad \text{for all } \lambda >0.
    \end{equation}
    Therefore, the left tail probability of random variable $Z_n$ satisfies
    \begin{equation}\label{eq:e29upper}
    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{-\lambda Z_n}}}{e^{-\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{-\lambda})}+ \lambda \delta }}^n,\quad \text{for all } \lambda >0,
    \end{equation}
    where the second equality follows from the moment generating function of $Z_n$,
    \[\bbE\off{e^{-\lambda Z_n}} = \of{1 - \alpha + \alpha e^{-\lambda}}^n. \]
    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = -\log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition. Plugging the minimizer into equation~\eqref{eq:e29upper}, we obtain the result
    \[ \bbP\off{Z_n \leq \delta n}  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha).  \]

@@

\begin{cor}[Sharp upper bounds on binomial right tails] Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in [1/2,1)$. Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$. The left tail probability of $Z_n$ is upper bounded as
	\begin{equation}\label{eq:lembinupper}
		\bbP\off{Z_n \geq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (\alpha,1),
	\end{equation}
	where $ D\of{\delta \Vert \alpha}$ is the KL divergence defined in Lemma~\ref{lem:sharpbin}.
\end{cor}

\begin{proof}
	Applying the Chernoff inequality directly to the random variable $Z_n$, we have
	 \begin{equation}
    	\bbP[Z_n \geq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda >0.
    \end{equation}

    Plugging the minimizer $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ into the above inequality, we obtain the results
    \[ \bbP[Z_n \geq \delta n]  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (\alpha,1).\]

@@

	\vspace{0.2cm}
	Since $Z_n$ is a binomial variable with size $n$ and probability $\alpha$, we have
	\begin{align}
		\bbP[Z_n \leq n\delta ] &= \sum_{k = 1}^m \begin{pmatrix}
			n \\ k
		\end{pmatrix} \alpha^k (1-\alpha)^{n - k}\\
		& \geq \begin{pmatrix}
			n \\ m
		\end{pmatrix} \alpha^m (1-\alpha)^{n - m}. \label{eq:e210pro1}
	\end{align}
	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1}, we obtain
	\[ \frac{1}{n} \log \bbP[Z_n \leq n\delta ] \geq \frac{1}{n} \log \begin{pmatrix}
			n \\ m		\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha). \]

@@

	\vspace{0.2cm}
	Plugging equation~\eqref{eq:e2102} into equation~\eqref{eq:e2101}, we have
	\begin{equation}\label{eq:e210pro3}
		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \phi(\tilde \delta) + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha) - \frac{\log(n+1)}{n} \geq -D(\delta \Vert \alpha ) - \frac{\log(n+1)}{n}.
	\end{equation}
	Multiplying by $n$ and take exponential on both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
	\begin{equation}
		\bbP \off{Z_n \leq n\delta} \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)}.
	\end{equation}
\end{proof}

@@

\subsection{Exercise 2.11}\label{Gausmax}
\begin{lem}[Gaussian maxima]
Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ normal random variables following $N(0,\sigma^2)$. Consider the random variable $Z_n \colonequals \max_{i\in[n]} |X_i|$. The expectation of $Z_n$ is upper bounded as follows
\begin{equation}\label{eq:e2111}
	\bbE[Z_n] \leq \sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}, \quad \text{for all } n\geq 2.
\end{equation}

The lower bound of $\bbE[Z_n] $ is
\begin{equation}\label{eq:e2112}
	\bbE[Z_n] \geq (1- 1/e) \sqrt{\sigma^2 \log n} , \quad \text{for all }n \geq 6.
\end{equation}
\end{lem}

The lower bound~\eqref{eq:e2112} is different to the original Exercise up to the constant.

@@

	\vspace{0.2cm}
	Consider the tail probability of $Z_n$. We have
	\begin{equation}
		\bbP[Z_n \geq t] = \bbP[\max_{i \in [n]} |X_i| \geq t] = 1 - \bbP[\max_{i \in [n]} |X_i| < t] = 1 - (1- \bbP[|X_1| \geq t])^n,
	\end{equation}
	where the last equality follows from the independence of $\{X_i\}_{i=1}^n$. By Bernoulli's inequality, we have
	\begin{equation}
		(1- \bbP[|X_1| \geq t])^n \geq 1 - n \bbP[|X_1| \geq t], \quad \text{for all } t >0.
	\end{equation} 
	Therefore, the expectation of $Z_n$ follows
	\begin{equation}\label{eq:e211pro1}
		\bbE[Z_n] = \int_{0}^{+\infty} \bbP[Z_n \geq t] dt \leq  c +  \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt, \quad \text{for all } c > 0.
	\end{equation}
	 Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the tail bound of $|X_1|$ satisfies
	 \begin{equation}\label{eq:e211proupper}
	 	\bbP[|X_1| \geq t] \leq 2 \sqrt{\frac{2}{\pi}} \frac{\sigma}{t} e^{-\frac{t^2}{2\sigma^2}}.
	 \end{equation}
	 Hence, let $ u = \frac{ t}{\sigma}$, the last integral in equation~\eqref{eq:e211pro1} is upper bounded as
	 \begin{align}
	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\  
	 	 %& = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} e^{-\frac{u^2}{2}} dt \\
	 	 & \leq \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} u e^{-\frac{u^2}{2}} dt \\
	 	 & = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}  e^{-\frac{c^2}{2\sigma^2}}. \label{eq:e211pro2}
	 \end{align}
	 For all $n \geq 2$, let $c = \sqrt{2\sigma^2 \log n}$ and plug the $c$ into equations~\eqref{eq:e211pro1} and \eqref{eq:e211pro2}. Then, we obtain the result
	 \begin{equation}
	 	\bbE[Z_n] \leq \sqrt{2\sigma^2 \log n} + \frac{2\sigma}{\sqrt{2\log n}}   \sqrt{\frac{2}{\pi}}  \leq \sqrt{2\sigma^2 \log n} + \frac{4\sigma}{\sqrt{2\log n}}, \quad \text{for all }n \geq 2.
	 \end{equation}

	 \vspace{0.2cm}
	 Next, we prove the equation~\eqref{eq:e2112}.

@@

	 The expectation of $Z_n$ is lower bounded as
	 \begin{equation}
	 	\bbE \off{Z_n} \geq \sigma \sqrt{\log n} \bbP \off{Z_n > \sigma \sqrt{\log n}}  = \sigma \sqrt{\log n} \off{1 - \of{1 - \bbP \off{|X_i| > \sigma \sqrt{\log n}}   }^n}.
	 \end{equation}

	 Since $|X_i|$ follows the half-normal distribution, the upper tail probability of $|X_i|$ is
	 \begin{equation}
	 	\bbP \off{ |X_i| > \sigma \sqrt{\log n}} = 1 - \text{erf}\of{\frac{\sqrt{\log n}}{\sqrt{2} } } \geq 1 - \sqrt{1 - n^{-\frac{2}{\pi}}},
	 \end{equation} 
	 where $\text{erf}(\cdot)$ is the CDF function of half-normal distribution, and the last inequality follows from the fact that $\text{erf}(x) \leq \sqrt{1 - \exp \of{\frac{4}{\pi} x^2 }}$. When $n \geq 6$, we have $1 - \sqrt{1 - n^{-\frac{2}{\pi}}} \geq \frac{1}{n}$, and thus the expectation satisfies
	 \begin{equation}
	 	\bbE \off{Z_n} \geq \sigma \sqrt{\log n} \off{1 - \of{1 - \frac{1}{n}   }^n} \geq \of{1 - \frac{1}{e}} \sigma \sqrt{\log n}, \quad \text{for all }n \geq 6.
	 \end{equation}

	 \iffalse
	 By Markov's inequality, the expectation of $Z_n$ satisfies
	 \begin{equation}
	 	\bbE[Z_n] \geq t \bbP[Z_n \geq t],\quad \text{for all }t \geq 0.
	 \end{equation}
	 By the independence of $\{X_i\}_{i=1}^n$, the tail probability of $Z_n$ is
	 \begin{equation}
	 	\bbP[Z_n \geq t] = 1 - (1 - \bbP(|X_1| \geq t))^n.
	 \end{equation}

@@

\begin{rmk}
	The absolute Gaussian maxima $Z_n$ concentrates around $\sigma \sqrt{\log n}$ for large $n$. Standard Gaussian variables concentrate around 0, while the maxima of the $n$ standard Gaussian variables is no longer a Gaussian variables. The scale of Gaussian maxima scales as $\sqrt{n}$, however, the sum of $n$ i.i.d.\ Gaussian variables scales as $n$.
\end{rmk}

@@

\begin{rmk}
	The maxima for $n$ i.i.d.\ sub-Gaussian variables also share the similar concentration property as Gaussian variables in Exercise~\ref{Gausmax}. The maxima $\max_{i \in [n]} X_i$ concentrates around the sub-Gaussian parameter $\sigma$.
\end{rmk}

@@

\begin{rmk}
	Without the independence assumption, the sum of sub-Gaussian variables is still a sub-Gaussian variable. Under the independence assumption, the product of two sub-Gaussian variables is a sub-exponential variable.
\end{rmk}

@@

	\iffalse
	\vspace{0.2cm}

@@

		\item[2.] Case 2: Suppose $t < 2\Delta$.
	\end{enumerate}

	\fi
	\end{proof}

	\begin{rmk}
		The concentration around mean implies the concentration around median for $n$ variables, vice versa. 
	\end{rmk}

@@

	\begin{rmk}
		Kernel density estimator satisfies the bounded difference property and possesses a concentration property around the mean.
	\end{rmk}

    \subsection{Exercise 2.16}
    \begin{lem}
    	Let $\{X_i\}_{i=1}^n$ be a sequence of independent variables taking values in a Hilbert space $\mathbb{H}$. Suppose that $\onorm{X_i}_{\mathbb{H}} \leq b_i$ almost surely, for all $i \in [n]$. Consider the real valued random variable $S_n = \onorm{\sum_{i=1}^n X_i}_{\mathbb{H}}$. The concentration bound of $S_n$ is
    	\begin{equation}\label{eq:e216a}
    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all } \delta \geq 0,
    	\end{equation}
    	where $b^2 = \frac{1}{n} \sum_{i=1}^n b_i^2$. The upper tail probability bound of $S_n$ is
    	\begin{equation}\label{eq:e216b}
    		\bbP \off{\frac{S_n}{n} \geq a + \delta} \leq e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0,
    	\end{equation}
    	where $a = \sqrt{\frac{1}{n^2} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
    \end{lem}

@@

    \begin{rmk}
    	The concentration inequalities also apply to the Hilbert space. 
    \end{rmk}

@@

    The proof for Hanson-Wright Inequality under the special case  $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.

@@

    	Consider $p_1$. For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and thus $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$. Given a constant $c$, $cX^2$ is still sub-exponential with parameters $(2 c, 4 c^2)$. Hence, for all $i \in [n]$, the variable $\mQ_{ii} X_i^2$ is sub-exponential with parameters $(2 \mQ_{ii}, 4 \mQ^2_{ii} )$. By the Bernstein-type inequalities (see Step 1 in reference), the probability $p_1$ is upper bounded as following,
    	\begin{equation}
    		p_1 = \bbP \off{\sum_{i=1}^n { \of{ \mQ_{ii}X^2_i - \mQ_{ii}\bbE[X^2_i]   } \geq \frac{t}{2} }} \leq \exp \of{ - \min \offf{  \frac{c t^2}{\sum_{i}^n \mQ^2_{ii} }, \frac{c't}{\max_{i \in [n]  }|\mQ_{ii}| } }},
    	\end{equation}
    	where $c$ and $c'$ are two constants independent with $n$ and $t$.

@@

    Reference: http://www-personal.umich.edu/~rudelson/papers/rv-Hanson-Wright.pdf.

    \begin{rmk}
    	For a random vector $X$, Hanson-Wright Inequality implies the concentration property for the matrix product $X^T \mQ X$, where $\mQ$ is a positive semi-definite matrix.
    \end{rmk}

@@

    \begin{rmk}
    	For a random variable $X$, the existence of $X$'s Orlicz norm implies the boundedness of the tail probability of $X$.
    \end{rmk}

@@

\begin{rmk}
	The maxima of Orlicz variables concentrates around the Orlicz norm $\sigma$ and scales as $\psi^{-1}(n)$. In previous exercises, the maxima of sub-Gaussian variables concentrates around the sub-Gaussian variable $\sigma$ and scales at $\log (n)$. 
\end{rmk}

@@

 \begin{rmk}
 	For $n$ independent variables $X_i$, the moment condition for $X_i$ implies the boundedness of the tail probability for the sample mean $\frac{1}{n} \sum_{i=1}^n X_i$.
 \end{rmk}

@@

 	 An alternative version of the proof under $\Delta_R > 0$ is below.
